--- 
title: "NYC-Citibike"
author: "Xiaonan Guo(xg2308), Gaoyi Shi (g32012), Kenny Jin (jj3134), Jiayin Lin (jl5802)"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

This project is inspired by Todd W. Schneider who is the Author of "A Tale of Twenty-Two Million Citi Bike Rides: Analyzing the NYC Bike Share System".[]https://toddwschneider.com/posts/a-tale-of-twenty-two-million-citi-bikes-analyzing-the-nyc-bike-share-system/#citibike-data

>"
Citi Bike is a privately owned public bicycle sharing system serving the New York City boroughs of the Bronx, Brooklyn, Manhattan, and Queens, as well as Jersey City, New Jersey. Named after lead sponsor Citigroup, it is operated by Motivate (formerly Alta Bicycle Share), with former Metropolitan Transportation Authority CEO Jay Walder as chief executive until September 30, 2018[1] when the company was acquired by Lyft. The system's bikes and stations use BIXI-branded technology from PBSC Urban Solutions.

>First proposed in 2008 by the New York City Department of Transportation, Citi Bike's scheduled 2011 opening was delayed by Hurricane Sandy and technological problems. It officially opened in May 2013 with 332 stations and 6,000 bikes. Annual expansions have brought the totals to 706 stations and 12,000 bikes as of October 2017, making the service the largest bike sharing program in the United States. Further expansions for Citi Bike are planned to extend its service area across the Bronx, Brooklyn, Manhattan, and Queens, and increase the number of bikes to 40,000."
>--[Wiki Link]https://en.wikipedia.org/wiki/Citi_Bike.

In 2020, due to the COVID 19 Pandemic, New York City implemented the lock down and social distance policy. Many Companies requires employees to work from home instead of going to office everyday. We want to analyze the pandemic's influence on New York City residents, specially for 1.6M Manhattan residents who use Citi Bike as part of their commute tools.

In our project, we will break up by different topics:


1. The Membership analysis (Age, Gender, Annual memberships or Subscriber)

2. Time Series on Bike Usage (2018 -2020)

3. Weekday/Weekend Usage

4. Weather and Temperature affects on the Bike usage

5. Bike users' routes pattern (How people use it as the commute tool)

6. The most popular stations and the impact of Covid 19 on the stations












<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources


The data source is taken from the official side of Citibike website.
Each file contains monthly trip data, which we take from 2018-01 to 2020-10, total 34 months.


[data source link:] https://s3.amazonaws.com/tripdata/index.html

There are 15 features in the dataset.

1. trip duration in seconds.

2. start time 

3. end time

4. start station id

5. start station name

6. start station latitude

7. start station longitude

8. end station id

9. end station name

10. end station latitude

11. end station longitude

12. bike id

13. user type

14. birth year

15. gender 

[You can use the R code to download our files:]  https://github.com/ds-eda-viz/NYC-Bike-Share/blob/main/sql/download_raw_data.R

<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

**Part 1 Basic exploration of variables **  
 -- continuous  
-- categorical  
Histogram  
Boxplot  
Bar chart  
....
You may not ride a Bicycle if You are under 16 years of age. A minor who is 16 years of age or older may only use the Services if the minor’s subscription is purchased by, and the minor is under the responsibility of, the minor’s parent or legal guardian.  
https://assets.citibikenyc.com/rental-agreement.html  
\newline

```{r}
library(dplyr)
library(mi)
library(ggplot2)
library(forcats)
```



```{r}
library(tidyverse)
data_201806 <- read.csv("rendered_data/201806-citibike-tripdata.csv",sep=",")
```

```{r}
data_201906 <- read.csv("rendered_data/201906-citibike-tripdata.csv",sep=",")
```

```{r}
data_202006 <- read.csv("rendered_data/202006-citibike-tripdata.csv",sep=",")
```

```{r}
str(data_201806)
```

```{r}
#Add Age column
data_201806$age=2020-data_201806$birth.year
```

```{r}
data_201806$year_month="201806"
```

```{r}
str(data_201806)
```
## Variable outliers  
We want to check tripduration and age  
\newline
Identifying outliers using plots and fitting models is relatively easy compared to what to do after identifying the outliers. Outliers can be rare cases, unusual values, or genuine errors. Genuine errors must be corrected if possible or else they must be removed.  
\newline

```{r}
ggplot(data_201806, aes(x=tripduration)) +
geom_boxplot(color = "darkgrey", fill = "lightBlue") +
ggtitle("201806-tripduration-distribution")
```

```{r}
# we created a column to categorize tripduration  
# website data source already removed duration less than 60 secs
data_201806$tripduration.bin<-sapply(data_201806$tripduration, function(x) 
                                           if (x<=3600) "less than 1h"
                                      else if (x<= 21600) "1h to 6h"
                                      else if (x<= 43200) "6h to 12h"
                                      else if (x<=86400) "12h to 24h"
                                      else if (x>86400) "more than 24h")
                                  
```

```{r}
str(data_201806)
```

```{r}
data_201806 %>%
  mutate(tripduration.bin=fct_relevel(tripduration.bin, c("less than 1h",
                                                          "1h to 6h",
                                                          "6h to 12h",
                                                          "12h to 24h",
                                                          "more than 24h"))) %>%
  ggplot(aes(x = tripduration.bin)) +
  geom_bar(color="black", fill = "lightblue") +
  ggtitle("201806-tripduration-distribution")
```

```{r}
ggplot(data_201806, aes(x=age)) +
geom_boxplot( fill = "lightBlue") +
ggtitle("201806-Age-distribution")
```


```{r}
ggplot(data_201806, aes(x=age)) +
geom_histogram(color="black", fill = "steelblue1", binwidth=5) +
ggtitle("201806-Age-distribution")
```


```{r}
ggplot(data_201806, aes(x=age, y=tripduration)) + 
  geom_point(color="blue") +
  ggtitle("Scatterplot for tripduration vs Age")

```

\newline
After checking data_201806,we did same analysis for data_201906, data_202006 as well

```{r}

```



**Part 2 Data Transformation **  

[Since each file contains more than Million rows, it's very slow to perform data transformation in R. We used Mysql server to process our data, all the SQL code have been used to generate datasets are located here:] https://github.com/xguo7013/NYC-Bike-Share/tree/main/sql



<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values




For simplicity of the analysis, we extracted abnormal data (missing data, outliers) using SQL and gathered the result into 2 csv files, abnormal_data.csv and abnormal_location.csv. To explore the missing value patterns, we only need abnormal_data.csv.

```{r}
library(dplyr)
library(mi)
library(ggplot2)
library(forcats)
```


```{r}
abnormal_data = read.csv("rendered_data/abnormal_data.csv", na.strings = "NULL")
```

## Missing values by Column and Row

```{r}
col_missing = colSums(is.na(abnormal_data))%>%sort(decreasing = TRUE)
col_missing
```

```{r}
df = data.frame(Column = fct_reorder(names(col_missing), col_missing), missing_count = col_missing)


ggplot(df, aes(x = Column, y = missing_count)) + ggtitle("Missing Count of Each Column") +
  geom_col(fill = "steelblue1") + coord_flip()
```

We can see that all missing entries are in 4 columns, the start station's name, id, and the end station's name id. The 4 columns have the __same__ missing value count of __2677__. In fact, it can be observed that if a row contains a missing value, all 4 columns (start station name, start station id, end station name, end station id) will contain missing values.

```{r}
row_missing = data.frame(table(rowSums(is.na(abnormal_data))))
names(row_missing) <- c('missing_count', 'row_count')
```

```{r}
row_missing
```

```{r}
ggplot(row_missing, aes(x = missing_count, y = row_count)) + ggtitle("Missing Count of Rows") + 
  geom_col(fill = "steelblue1")
```

We can see that there are __2677__ rows that contain 4 missing entries, and other rows don't have any missing entries. This verifies that if a row contains a missing value, all 4 columns (start station name, start station id, end station name, end station id) will be missing.

## Missing pattern by Year and Month

```{r}
monthly_missing = read.csv("rendered_data/monthly_missing.csv")
monthly_missing$month = as.factor(monthly_missing$month)
```


```{r}
ggplot(monthly_missing, aes(x=month, y=num_trips)) + geom_col(fill="steelblue1") + ggtitle("Missing Count of Each Month") + facet_grid(monthly_missing$year)
```

The plot above shows that there are many missing entries in 2018, especially 2018-08, with more than 1000 rows that have missing values. There are only a few missing rows in 2019, and there are no missing rows in 2020. 

## Dealing with missing values

We decided that we discard all these rows that contain missing values and outliers __before__ we conduct any further analysis. 

It should be noted that there are only 2677 rows that contains missing values. If we include rows that have outliers, we only need to discard about 50000 rows. Comparing to the overall data size of about 55 million rows, we are only throwing away a small portion, which will not largely impact our analysis.




<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results



## The Membership analysis (Age, Gender, Annual memberships or Subscriber)


1.Color Palette  
First, we carefully pick our color palette for plots in this project, to make it consistent with company brand. The base colors for Citibike is blue and red.  
Color patette chosen are steelblue series and indianred series.  

```{r}
Palette=c('steelblue2','indianred1') 

```


2.Age/gender vs bike usage 
We care about how gender and age affect the bike usage, what's the difference betwwen each group and what kind of patterns there. Also, we want to see if there are any changes among years. Two facet line charts were created based on total number of trips and average duration of trips.  
Three features were considered, age, gender and year  

```{r}
library(tidyverse)
```

```{r}
df1 <- read.csv("rendered_data/age_gender_duration.csv")
df1$gender <- as.factor(df1$gender)
df1<- df1 %>% mutate(gender= fct_recode(gender,"unknown"="0","male"="1","female"="2"))%>%
      filter(gender!='unknown')

```


```{r,fig.width=10, fig.height=4}
df1 %>%
  ggplot(aes(x=age, y=trip_count/1000,group=gender))+
  geom_line(aes(color=gender))+
  geom_point(aes(color=gender))+
  scale_x_continuous(breaks=seq(0,100,10))+
  scale_color_manual(values=Palette)+
  ylab("trip_count (/1000)")+
  facet_wrap(~year)+
  ggtitle("NYC Citibike Total Number of Trips by Age and Gender" )
  
```

```{r,fig.width=10, fig.height=4}
df1 %>%
  ggplot(aes(x=age, y=trip_duration_avg,group=gender))+
  geom_line(aes(color=gender))+
  geom_point(aes(color=gender))+
  scale_x_continuous(breaks=seq(0,100,10))+
  scale_color_manual(values=Palette)+
  facet_wrap(~year)+
  ylab("trip_duration_avg(minites)")+
  ggtitle("NYC Citibike Average Trip Duration by Age and Gender" )
```

Note: For 2020, data was collected from January to October  

We can clearly see some patterns from above two plots.  
1) Males made more trips than females amnong almost all age levels. Without checking the accurate number, it looks like males have more than twice of the total trips made by females  
2) Total trips increases before age 30, but decreases after that, both males and females have total trip peaks around age 30. Young people tend to use bikes more.  
3) In terms of average duration,females tend to have longer trip duration then males. It makes sense as in general, makes ride faster then females.  
4) An interesting thing is that below age 65, the average duration is very stable for both male and female, between 10 and 20 minutes, but after age 65, it varies much. We can think about people may use bike a lot for daily commuting purpose on Mahattan. But after retirement, people bike for different reasons, maybe for exercise or for short trips. 
5) Another thing is that the average duration in 2020 seems longer then previous years. Probably people who need to go outside reduce the use of public transportation, and ride bikes more to get to destination
6) Finally, we think in general there is slight but not that significant pattern changes in year 2020 based on age and gender.  



3.Membership types vs bike usage
We want to look into more details to check if user behavior differs among age, gender and year. There are two Membership types, pass user and subscriber. We want to see which one peple use more. Two plots made here, one is stacked bar chart to show the total numbers, another one is mosaic plot to understand the association among variables, mainly gender and age in diferent years. 

Ages were binned into categories for analysis.


```{r}
df2 <- read.csv('rendered_data/usertype_gender_age.csv')
df3<-df2
df3$gender <- as.factor(df3$gender)
df3<- df3 %>% mutate(gender= fct_recode(gender,"unknown"="0","male"="1","female"="2"))

df4<-df2 %>% filter(gender!=0)
df4$gender <- as.factor(df4$gender)
df4<- df4 %>% mutate(gender= fct_recode(gender,"male"="1","female"="2"))
```




```{r,fig.width=9, fig.height=4}
Palette2=c('indianred1','steelblue2') #indianred1,lightcoral,salmon2,coral1
df3 %>%
  ggplot(aes(x=age_bin, y=total_trip_count/1000, fill=usertype))+
  geom_col()+    #geom_col(position = "dodge")
  scale_fill_manual(values=Palette2)+
  facet_wrap(~year)+
  ggtitle("NYC Citibike Membership type by Year, Age")

```

Note: 
1) Customer = 24-hour pass or 3-day pass user; Subscriber = Annual Member  
2) For 2020, data was collected from January to October  



We have some findings below.  
1) We found in both bar chart, for different age bins, age 45-54 tend to have the highest percentage of customer type, followed by age below 35 which have relatively higher percentage of customer type. While age 35-44 users and age higher then 55 users tend to have more subscribers.
2) We have some doubts for age 45-54, and curious why this age bin have the highest customer type. This question may be answered by mosaic plot.
3) Another thing we can tell from the plots is that due to covid-19, the percentage of customer increased, people work from home and stay at home may cancel the annual membership. But in general, the usage doesn't drop largely.  


```{r}
library(vcd)

df_2019<-df3%>%filter(year==2019)

Palette3= c('grey','steelblue1','steelblue')
mosaic(xtabs(total_trip_count~ gender+age_bin+usertype , df_2019),
             direction = c("h", "v", "h"),
             gp = gpar(fill = Palette3),
             main = "Membership type by Gender, Age in 2019")
```
An interesting thing is that for gender 'unknown', age 45-54 occupies the lagest portion and membership type 'customer' is the main type in this bin, which is very different from other catergories. We cannot find a good reason for this abnomality, and guess when people create new profile, they don't want to input the true information and keep the default settings, especially for membership type 'customer', who care more about to use bike as soon as possible.  
For this strange pattern in unknown category, we will focus more on male and female categories, and filter the data for gender unknown for further analysis.



```{r,fig.width=9, fig.height=4}
library(vcd)

df_2019<-df4%>%filter(year==2019)
df_2020<-df4%>%filter(year==2020)

Palette3= c('steelblue1','steelblue')
mosaic(xtabs(total_trip_count~ gender+age_bin+usertype , df_2019),
             direction = c("h", "v", "h"),
             gp = gpar(fill = Palette3),
       main = "Membership Types by Gender, Age in 2019")

mosaic(xtabs(total_trip_count~ gender+age_bin+usertype , df_2020),
             direction = c("h", "v", "h"),
             gp = gpar(fill = Palette3),
       main = "Membership Types by Gender, Age in 2020")


```
Note: 
1) The plot should show the fill in color for the last cut - usertype, but it's not working here.  
2) In order to show more clear plot, we made the first cut horizontally  


Mosaic plot mainly show the proportion and association for gender, age and usertypes. The patterns are now more clear then stacked bar chart.  
1) For both male and female, yonger ages tend to have more customer type, while elder people tend to have more subscribers. We guess elder people may have more fixed pace of life and commute between home and workplace daily; and for age higher than 55, those people may bicycle for exercise and healthier life style.  
2) We don't see much difference for usertypes in terms of age and gender.  

2) Same thing we can tell from mosaic plot, due to covid-19, the percentage of customer increased  




## Time Series on Bike Usage (2018 -2020)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
```

Question 1: How does number of trips differ by years from 2018 to 2020? How is it affected by COVID-19? Is there any trend? 

From the time series line graph, we can see that the trends in 2018 and 2019 are quite similar. Both of them started to increase in Feburary, reached a relatively high level from May to October, and started to decrease as winter came. In 2020, due to COVID-19, there was a sharp decrease in April 2020. However, surprisingly, the number of trips began to increase right after April, and followed the same pattern as it was in previous years. Based on COVID-19 safety instructions, people should avoid outdoor activities and interaction with people, but it seems that citizens in NYC are not following these instructions, and that is why COVID-19 is still a big problem in NYC, or even in the entire US. 

```{r}
df_1 <- read.csv("rendered_data/monthly_usage.csv")

df_1[,'num_trips'] = df_1[,'num_trips']/1000
df_1[,'year']<-factor(df_1[,'year'])
df_1[,'month']<-factor(df_1[,'month'])

```

```{r,fig.width=9, fig.height=4}
ggplot(df_1, aes(x=month, y = num_trips, group = year, color = year)) + 
  geom_line(size =1) + 
  ggtitle("Number of Trips with Different years (In Thousands)") + 
  xlab("Month") + 
  ylab("Number of Trips")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))
```

```{r,fig.width=9, fig.height=4}
ggplot(df_1, aes(x=month, y = num_trips, group = year, color = year)) + 
  geom_line(size =1) + 
  facet_grid(rows = vars(year)) + 
  ggtitle("Number of Trips with Different years (In Thousands)") + 
  xlab("Month") + ylab("Trips")+
  ylab("Number of Trips")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))
```



## Weekday/Weekend Usage

Can we extract any pattern for trips during weekdays and weekends? What time in a day are bikes used most frequently?

Here we created bar chart analysis for hourly bike usage by weekdays and weekends. We can see that during weekdays, there are 2 peeks. In the morning, the peek occurs at 8 o'clock, when people are busy getting to their workplace. In the afternoon, the peek shows up at 17 and 18 o'clock, when people just end their work and start to get back home. During weekends, we can see that the number of trips almost follows a normal distribution, where it reaches its peek at 15 o'clock. Such trend occurs probably because most people are using Citi bikes for sightseeing during weekends. Combined with other data (for example, data of user info), we might be able to determine the proportion of people that mainly use Citi bikes for work, and thus modifying the bike deployment or the location of bike stations accordingly. 
```{r}
df_2 = read.csv("rendered_data/weekend_weekday_dist.csv")

```

```{r}
df_2[,'total_num_trips'] = df_2[,'total_num_trips']/1000
df_2[,'is_weekend']<-factor(df_2[,'is_weekend'])
df_2[,'hour']<-factor(df_2[,'hour'])
```

```{r,fig.width=9, fig.height=4}
ggplot(df_2, aes(x=hour, y = total_num_trips, group = is_weekend)) + geom_bar(stat="identity", fill = "steelblue2") + facet_grid(rows = vars(is_weekend)) + ggtitle("Number of Trips by Hour (in 1000)") + xlab("Hour") + ylab("Trips")
```




## Temerature affects on the Bike usage  
```{r}
library(ggplot2)
```

```{r}
df =read.csv('rendered_data/monthly_usage_all.csv')
```

In this section, we will analyze the bike usage from temperature perspective. Trips are bellow 1 Million if the temperature bellow 45 degrees in 2018 and 2019, then every temperature goes up 10 degrees, the average usage goes up 0.5M. Only 2020 is an exception, we see a huge decrease from 45 - 55 degrees in 2020. I have did the analysis, usually March, April and May as average monthly temperature between 45 - 55. During this time, New York City takes very strict lock down policy. The New York State governor closed all the restaurants, gym, public parks and sports fields to prevent the spread of corona virus.
Then the ride usage bounce back on Jun and July since the lock down policy has been released. With the summer approaching, the usage backs to normal level as 2018 and 2019.

From the analysis, we can the best weather to take a ride would be 65 to 75 range, there are around 2Million people would choose Citibike over taking cars or subways in New York City. Let's go to take a ride on the Citibik when the weather is nice!

```{r}
df_2018 <-df[df$month<8,]


ggplot(data =df_2018, aes(x=month, y=num_trips/1000000))+
  geom_line(color ='steelblue2',size =1)+ facet_wrap(~year) +
  scale_x_discrete(limits=c("25","35","45","55","65","75","85"))+
  ggtitle(" NYC Citi Bike Usage with temperature ", 
          "Based on Citi Bike system data,  2018 -2020 data")+
  labs(x ="Temperature (F)", y ="Trips in Millions")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))
```




## Bike Users' route pattern(How people use it as the commute tool)


```{r}
#install this library if you do not have it
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/ggpubr")

#install.packages("ggpubr")



####install package for ploting graphs in one figure

```
1. First, we took the data from Citibike official website, there are station location which can be mapped to different counties, we picked all the stations in Manhattan, Queens, Brooklyn as our main target to analyze the route pattern for riders. There are over 1.6 Millions residents in Manhattan, we can see 8 am - 10 am and 5 pm - 7pm are the peak hours for bike usages, average bike trips are over 100K times.It shows there are around 10% of Manhattan residents are using Citibike for their commute.2018 and 2019 has the same pattern, but 2018 the trip numbers drops to around 50K in morning rush hours, this is largely caused by COVID -19 pandemic, which companies require employee to work from home and the city implement lock down policy starts from end of March 2020 to July 2020.
Majority of bike trips happen within the Manhattan, we want to analyze the cross-borough bike usage, we need to drop the Manhattan trips and analyze the Manhattan to Queens and Manhattan to Brooklyn trips, which will display in the next plot. 

```{r}

library(ggplot2)
library(ggpubr)

#use the above code to install ggpubr if you do not have it

```


```{r}
##change directory
##read data files



df_2018 =read.csv('rendered_data/borough_trips_201806.csv',sep =',')

df_2019 =read.csv('rendered_data/borough_trips_201906.csv',sep =',')

df_2020 =read.csv('rendered_data/borough_trips_202006.csv',sep =',')

```


```{r}
###Mahanttan data

man_2018 =df_2018[df_2018$start_borough=="Manhattan", ]

man_2018_plot <- ggplot(data = man_2018, 
                        aes(x=hour,
                        y=total_num_trips/1000, 
                        color =route))+
  geom_point()+geom_line() +
  ggtitle("2018 NYC Citi Bike Trips in Manhattan ", 
          "Based on Citi Bike system data, 2018-06-01 to 2018-06-30") +
  labs(x ="Trip Hours", y ="")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

###2019
man_2019 =df_2019[df_2019$start_borough=="Manhattan", ]

man_2019_plot <- ggplot(data = man_2019, 
                        aes(x=hour,
                            y =total_num_trips/1000, 
                            color =route))+
  geom_point()+geom_line() +
  ggtitle("2019 NYC Citi Bike Trips in Manhattan ", 
          "Based on Citi Bike system data,  2019-06-01 to 2019-06-30") +
  labs(x ="Trip Hours", y ="Average Hourly Citi Bike trips(Thousands)")+
  coord_cartesian(ylim = c(0, 150))+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'),
        axis.title.y = element_text(size = 9))
   


###2020
man_2020 =df_2020[df_2020$start_borough=="Manhattan", ]

man_2020<-as.data.frame(man_2020)

man_2020_plot <-  ggplot(data = man_2020, 
                         aes(x=hour,
                             y =total_num_trips/1000, 
                             color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 150))+
  ggtitle("2020 NYC Citi Bike Trips in Manhattan ", 
          "Based on Citi Bike system data,  2020-06-01 to 2020-06-30") +
  labs(x ="Trip Hours", y ="")+
  theme_grey(9)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))


```

```{r,fig.width=9, fig.height=4}


ggarrange(man_2018_plot, man_2019_plot, man_2020_plot, 
          label.y = 1,
          align='hv',
          #common.legend = TRUE,
          ncol = 1, nrow = 3)
```


```{r}

##Manhattan to Outers

man_outer_2018 =df_2018[df_2018$start_borough=="Manhattan" & df_2018$end_borough!="Manhattan", ]

man_outer_2018_plot <- ggplot(data = man_outer_2018, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +
  ggtitle("2018 NYC Citi Bike Trips from Manhattan to outers ", 
          "Based on Citi Bike system data,  2018-06-01 to 2018-06-30") +
  labs(x ="Trip Hours", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

## 2019

man_outer_2019 =df_2019[df_2019$start_borough=="Manhattan" &df_2019$end_borough!='Manhattan', ]

man_outer_2019_plot <- ggplot(data = man_outer_2019, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +
  ggtitle("2019 NYC Citi Bike Trips from Manhattan to outers ", 
          "Based on Citi Bike system data,  2019-06-01 to 2019-06-30") +
  labs(x ="Trip Hours", y ="Average Hourly Citi Bike trips")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'),
        axis.title.y = element_text(size = 9))

##2020

man_outer_2020 =df_2020[df_2020$start_borough=="Manhattan"&df_2020$end_borough!='Manhattan', ]

man_outer_2020 <-as.data.frame(man_outer_2020)

man_outer_2020_plot <-  ggplot(data = man_outer_2020, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 7000))+
  ggtitle("2020 NYC Citi Bike Trips from Manhattan to outers ", 
          "Based on Citi Bike system data,  2020-06-01 to 2020-06-30") +
  labs(x ="Trip Hours", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

```

2. Manhattan to Outers
From the pattern we can see, in the morning few people will take the bikes but in the afternoon, there are average 6K trips each day. People tend to take other transformations for going to queens and Brooklyn and coming back to Manhattan with bikes. 
One interesting point is the morning bike usage doubled in 2020, the average bike usage around 10AM is 1k in both 2018 and 2019 but it is 2K in 2020. This can be implied that people try to take social distance and avoid taking subway or bus for their commute and choose to ride bikes instead.

```{r,fig.width=9, fig.height=4}
##Manhattan to outers
ggarrange(man_outer_2018_plot, man_outer_2019_plot, man_outer_2020_plot, 
          label.y = 1,
          align='hv',
          ncol = 1, nrow = 3)

```


```{r}
####Queens to outer

###2018

qn_outer_2018 =df_2018[df_2018$start_borough=="Queens" & df_2018$end_borough!="Queens", ]

qn_outer_2018_plot <- ggplot(data = qn_outer_2018, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 1500))+
  ggtitle("2018 NYC Citi Bike Trips from Queens to outers ", 
          "Based on Citi Bike system data,  2018-06-01 to 2018-06-30") +
  labs(x ="Trip Hours", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

###2019

qn_outer_2019 =df_2019[df_2019$start_borough=="Queens" &df_2019$end_borough!='Queens', ]

qn_outer_2019_plot <- ggplot(data = qn_outer_2019, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 1500))+
  ggtitle("2019 NYC Citi Bike Trips from Queens to outers ", 
          "Based on Citi Bike system data,  2019-06-01 to 2019-06-30") +
  labs(x ="Trip Hours", y ="Average Hourly Citi Bike trips")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'),
        axis.title.y = element_text(size = 9))


qn_outer_2020 <- df_2020[df_2020$start_borough=="Queens"&df_2020$end_borough!='Queens', ]

qn_outer_2020 <- as.data.frame(qn_outer_2020)

qn_outer_2020_plot <-  ggplot(data = qn_outer_2020, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 1500))+
  ggtitle("2020 NYC Citi Bike Trips from Queens to outers ", 
          "Based on Citi Bike system data,  2020-06-01 to 2020-06-30") +
  labs(x ="Trip Hours", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

```

3. Queens to Outer
Very few people take Citi bikes in Queens since there are fewer bike stations in Queens. It still worths us to take a look on riders' pattern.
In the morning rush hours, there are around 700 riders take bikes as their way to Manhattan and 200 riders to Brooklyn for both 2018 and 2019. Things change a little bit in 2020, there are around 1000 riders take bikes to Manhattan, 400 riders to Brooklyn. Rider numbers almost doubled in 2020. It has the similar pattern as people live in Manhattan and go to Queens or Brooklyn in the morning. Due to pandemic and social distance advocate, people in New York city tend to ride more bikes than usual to avoid taking subways or buses in the morning rush hours.

```{r,fig.width=9, fig.height=4}
###Queens to Outers

ggarrange(qn_outer_2018_plot, qn_outer_2019_plot, qn_outer_2020_plot, 
          label.y = 1,
          align='hv',
          ncol = 1, nrow = 3)
```




```{r}
##Brooklyn to outers

bk_outer_2018 =df_2018[df_2018$start_borough=="Brooklyn" & df_2018$end_borough!="Brooklyn", ]

bk_outer_2018_plot <- ggplot(data = bk_outer_2018, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +
  ggtitle("2018 NYC Citi Bike Trips from Brooklyn to outers ", 
          "Based on Citi Bike system data,  2018-06-01 to 2018-06-30") +
  labs(x ="Trip Hours", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

bk_outer_2019 =df_2019[df_2019$start_borough=="Brooklyn" &df_2019$end_borough!='Brooklyn', ]

bk_outer_2019_plot <- ggplot(data = bk_outer_2019, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 6000))+
  ggtitle("2019 NYC Citi Bike Trips from Brooklyn to outers ", 
          "Based on Citi Bike system data,  2019-06-01 to 2019-06-30") +
 labs(x ="Trip Hours", y ="Average Hourly Citi Bike trips")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'),
        axis.title.y = element_text(size = 9))

bk_outer_2020 =df_2020[df_2020$start_borough=="Brooklyn"&df_2020$end_borough!='Brooklyn', ]

bk_outer_2020 <-as.data.frame(bk_outer_2020)

bk_outer_2020_plot <-  ggplot(data = bk_outer_2020, aes(x=hour,y =total_num_trips, color =route))+
  geom_point()+geom_line() +coord_cartesian(ylim = c(0, 6000))+
  ggtitle("2020 NYC Citi Bike Trips from Brooklyn to outers ", 
          "Based on Citi Bike system data,  2020-06-01 to 2020-06-30") +
  labs(x ="Trip Hours", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))


```

4. Brooklyn to Outers
Brooklyn has a different story compare with their peers in Manhattan and Queens. In 2018 and 2019, more people take bikes to go to other counties with average 6500 trips in the morning rush hours, it dropped to 3000 trips in 2020. I performed a deeper analysis on Brooklyn residents. Most of those trips between Brooklyn bridge and lower Manhattan, since a lot of financial companies are taking strict work from home policy. It explains why riders reduce so many between Brooklyn and Manhattan but not many changes between Brooklyn and Queens.

```{r,fig.width=9, fig.height=4}

###Brooklyn to Outers
ggarrange(bk_outer_2018_plot, bk_outer_2019_plot, bk_outer_2020_plot, 
          label.y = 1,
          align='hv',
          ncol = 1, nrow = 3)
```


## Top Popular stations by year

```{r}

st_2018 <- read.csv('rendered_data/top_10_stations_2018.csv')

st_2019 <- read.csv('rendered_data/top_10_stations_2019.csv')

st_2020 <- read.csv('rendered_data/top_10_stations_2020.csv')

 plt_1 <- ggplot(st_2018, aes(trips_in_thousands, reorder(station_name, trips)))+ geom_point()+
  ggtitle("2018 NYC Citi Bike popular station", 
          "Based on Citi Bike system data,  2018-01-01 to 2018-12-31") +
  labs(x ="Trip in Thousands", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

plt_2 <- ggplot(st_2019, aes(trips_in_thousands, reorder(station_name, trips)))+ geom_point()+
  ggtitle("2019 NYC Citi Bike popular station", 
          "Based on Citi Bike system data,  2019-01-01 to 2019-12-31") +
  labs(x ="Trip in Thousands", y =" Station Names ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))


plt_3 <- ggplot(st_2020, aes(trips_in_thousands, reorder(station_name, trips)))+ geom_point()+
  ggtitle("2020 NYC Citi Bike popular station", 
          "Based on Citi Bike system data,  2020-01-01 to 2020-12-31") +
  labs(x ="Trip in Thousands", y =" ")+
  theme_grey(10)+
  theme(plot.title =element_text(face='bold', color ='steelblue2'))

```

In this topic, we want to get the most popular stations among 3 years, and see the impact of the Covid 19 on the bike usage in stations.

In both 2018 and 2019, the Pershing Square North station is the most population for bike users.
"
Pershing Square is a public square in Manhattan, New York City, located where Park Avenue and 42nd Street intersect in front of Grand Central Terminal. The main roadway of Park Avenue crosses over 42nd Street on the Park Avenue Viaduct, also known as the Pershing Square Viaduct. Two service roads, one northbound and one southbound, connect 42nd Street with the main roadway of Park Avenue, at 40th Street.

Pershing Square was named after John J. Pershing in 1919. The name was originally supposed to apply to the block bounded by Park Avenue, Lexington Avenue, 41st Street, and 42nd Street. Three buildings were ultimately developed on the block in the 1920s: the Pershing Square Building, 110 East 42nd Street, and the Chanin Building. Subsequently, the name applied to the service roads of the Park Avenue Viaduct. A tourist information center under the viaduct, at Pershing Square, was built in 1939; it was later reconfigured to be a store and then a restaurant. The service roads between 42nd and 41st Streets were converted into a pedestrian public plaza in 2018.
"[wiki:]https://en.wikipedia.org/wiki/Pershing_Square,_Manhattan

This is one of the most interesting place in Manhattan, tourist might contribute a lot in the bike usage.






```{r}

plt_1

```


```{r}

plt_2

```

2020 is a different story, Pershing Square North is not longer the top one, it's even not the top3.
it falls to 11th position, this is largely impact by the covid-19 which New York City takes lock down policy and people choose to stay home, the number of tourists decreased significantly in this year. 
The top 12 Ave & W40 Bike station is in the terminal of New York and New Jersery Ferry. A lot of people might take the ferry then use the bike as their commute tool inside Manhattan instead of taking subway.

```{r}

plt_3


```

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component


<!DOCTYPE html>

<html lang="en">
    
    <head>
        <meta charset="utf-8">
        <title></title>
        <script src="https://d3js.org/d3.v6.js"></script>
        <style type ="text/css">
        </style>
    </head>
    
    <body>
    
    <svg width="1000" height="300" id="svg1"></svg>
    <svg width="1000" height="300" id="svg2"></svg>
    
    <script>
    
      var svg1 = d3.select("#svg1")
      var margin = {top: 20, right: 50, bottom: 30, left: 50}
      var width =  svg1.attr("width") - margin.left - margin.right
      var height = svg1.attr("height") - margin.top - margin.bottom
      var g1 = svg1.append("g").attr("transform", `translate(${margin.left}, ${margin.top})`);

      var xScale = d3.scaleBand().domain(d3.range(24)).range([0, width]).paddingInner(0.05);
      var yScale = d3.scaleLinear().domain([0, 4000]).range([height, 0]);
        
      var xAxis = d3.axisBottom(xScale);
      var yAxis = d3.axisLeft(yScale);
      
      g1.append("g").attr("transform", `translate(0, ${height})`).call(xAxis);
      g1.append("g").call(yAxis);
      g1.append("g").call(yAxis);
      
      var readfile = function(d){
        return {
          is_weekend: d.is_weekend,
          hour: d.hour,
          trips: d.total_num_trips/1000 
        }
      };
      
      d3.csv("https://raw.githubusercontent.com/ds-eda-viz/NYC-Bike-Share/main/rendered_data/weekend_weekday_dist.csv", readfile)
        .then(function(data){
          
          var weekday = data.filter(function(d) {return d.is_weekend == "weekday";});
          var bars = g1.selectAll(".bar").data(weekday).enter().append("rect").attr("class", "bar")
          
          bars.attr("x", function(d) {return xScale(d.hour);})
          .attr("y", function(d) {return yScale(d.trips);})
          .attr("width", xScale.bandwidth()*0.95)
          .attr("height", function(d) {return height - yScale(d.trips)})
          .attr("fill", "steelblue")
          .on("mouseover", function(){ d3.select(this).attr("fill", "orange");})
          .on("mouseout", function(){ d3.select(this).attr("fill", "steelblue");});
          
          g1.append("text")
          .attr("x", "10")
          .attr("y", "20")
          .text("Number of Trips during Weekdays");
          g1.append("text")
          .attr("x", "10")
          .attr("y", "40")
          .text("(in thousands)");
          
        });
        
      /////////////
      
      var svg = d3.select("#svg2")
      var margin = {top: 20, right: 50, bottom: 30, left: 50}
      var width =  svg.attr("width") - margin.left - margin.right
      var height = svg.attr("height") - margin.top - margin.bottom
      var g2 = svg.append("g").attr("transform", `translate(${margin.left}, ${margin.top})`);
      
      var xScale = d3.scaleBand().domain(d3.range(24)).range([0, width]);
      var yScale = d3.scaleLinear().domain([0, 4000]).range([height, 0]);
        
      var xAxis = d3.axisBottom(xScale);
      var yAxis = d3.axisLeft(yScale);
      
      g2.append("g").attr("transform", `translate(0, ${height})`).call(xAxis);
      g2.append("g").call(yAxis)
      
      d3.csv("https://raw.githubusercontent.com/ds-eda-viz/NYC-Bike-Share/main/rendered_data/weekend_weekday_dist.csv", readfile)
        .then(function(data){
          var weekend = data.filter(function(d) {return d.is_weekend == "weekend";})
          var bars = g2.selectAll(".bar").data(weekend).enter().append("rect").attr("class", "bar")
          
          bars.attr("x", function(d) {return xScale(d.hour);})
          .attr("y", function(d) {return yScale(d.trips);})
          .attr("width", xScale.bandwidth()*0.95)
          .attr("height", function(d) {return height - yScale(d.trips)})
          .attr("fill", "steelblue")
          .on("mouseover", function(){ d3.select(this).attr("fill", "orange");})
          .on("mouseout", function(){ d3.select(this).attr("fill", "steelblue");});
          
          g2.append("text")
          .attr("x", "10")
          .attr("y", "20")
          .text("Number of Trips during Weekends");
          g2.append("text")
          .attr("x", "10")
          .attr("y", "40")
          .text("(in thousands)");
        });
    
    </script>
    
    </body>

</html>








<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion

Everyday we see people ride bikes pass us and take tours in New York City, but few of us spend time on studying this system. By using the R and D3 programming in our exploratory data analysis class. Four of us are able to perform a deep analysis on the bike usage daily, monthly and yearly. There are a lot of interesting part for this system.

1. Male riders are more then female riders among all age levels, people between 30 - 35 groups are riding bikes more than other 
groups. Although female rides has less trips, average trip duration is longer than male riders. In 2020, trip duration is longer 
than average duration in 2018, 2019. More people would rely on the bike as their commute tools in this year.

2. In a normal year, bike usage would be increased from March to September, but the lock down in April and May 2020 caused the 
readerships reduced significantly. After the lock down which is June, we see the ridership has a huge amount of usage this year.
People might do not have confidence to take Subway and Bus, so they choose to ride the bike for commuting.

3. The route pattern is another interesting part we want to point out. In 2018 and 2019, few people would ride bikes to cross 
boroughs. We see an increase in 2020, more and more people would ride bikes from Brooklyn or Queens to Manhattan. As the social 
distance policy implements in New York City, people tend to take bikes for their work. 

4. By analyzing the most popular bike stations, we come with a conclusion that people do not travel to New York City this year.
As the center of the pandemic, New York City has closed restaurants, gym, public libraries, movie theaters and shopping stores. 
Both retail and travel industry would suffer a huge loss from this year due to the pandemic. 

Very soon, there will be single digit number of days left for the year of 2020, People who live in this city would never forget 
about this year. We will remember how this pandemic changed our lives forever, and friends and families we have lost forever. We 
believe our city will recover from it and life will be back to normal. It's maybe a long wait, but we believe it.





<!--chapter:end:08-conclusion.Rmd-->

